---
title: "Reproducible research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download, unzip, and read data

```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
dest <- "activity.zip"
download.file(fileUrl, destfile=dest, mode='wb')
unzip(dest)
activity <- read.csv("activity.csv")
```

## Histogram of the total number of steps taken each day
```{r}
library(dplyr)
library(ggplot2)
Nsteps <- activity %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) 
g <- ggplot(Nsteps, aes(x=steps)) + geom_histogram(binwidth=500) + xlab("Steps") + ylab("Count") + ggtitle("Total number of steps taken per day")
g
```

## Mean and median number of steps taken each day
```{r}
meanstep <- mean(Nsteps$steps, na.rm=TRUE)
medianstep <- median(Nsteps$steps, na.rm=TRUE)
print("Mean number of steps taken each day : ")
meanstep
print("Median number of steps taken each day : ")
medianstep
```

## Time series plot of the average number of steps taken
```{r}
Msteps <- activity %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(steps = mean(steps))
with(Msteps, plot(interval, steps, type='l', xlab="Interval", ylab="Steps", main="Average number of steps taken"))
```

## 5-minute interval that, on average, contains the maximum number of steps
```{r}
maxinterval <- Msteps[which.max(Msteps$steps),1]
maxinterval
```

## Impute missing data
### Total number of missing values in the dataset
```{r}
sum(is.na(activity$steps))
```
### Strategy for filling in all of the missing values in the dataset
### I will try to substitute all the NA's with the mean value for the corresponding intervals

## Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity.na.rm <- activity
na <- is.na(activity.na.rm$steps)
avgint <- tapply(activity.na.rm$steps, activity.na.rm$interval, mean, na.rm=TRUE)
activity.na.rm$steps[na] <- avgint[as.character(activity.na.rm$interval[na])]
```

##Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
nomissing <- activity.na.rm %>%
  group_by(date) %>%
  summarize(steps = sum(steps))

ggplot(nomissing, aes(x=steps)) + geom_histogram(binwidth=500)
meanstep <- mean(nomissing$steps)
medianstep <- median(nomissing$steps)
```


##Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
library(lattice)
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
activity.na.rm$day <- as.factor(ifelse(weekdays(as.Date(activity.na.rm$date)) %in% weekdays, "Weekday", "Weekend"))
steps_day <- aggregate(steps~interval+day, activity.na.rm, mean)
with(steps_day, xyplot(steps~interval|day, type="l", main="Average steps per day (weekday vs weekend)"))
```